Dissertation Proposal

### 2023-11-08 proposal.tex
disable line numbers
grammar check

\usepackage[switch]{lineno}
\usepackage{xcolor}
\renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
\usepackage[utf8]{inputenc} % Use UTF-8 encoding
\usepackage[T1]{fontenc} % Use the T1 font encoding
\usepackage{geometry} % Adjust page margins
\usepackage{amsmath, amssymb} % For math symbols and equations
\usepackage{graphicx} % For including images
\usepackage{titlesec} % For customizing section titles
% \usepackage[style=authoryear]{biblatex}  % For managing references
\usepackage[round, square]{natbib}
\bibliographystyle{plainnat}
\usepackage{hyperref}
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{tabularx}   %% For tables with growable column
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\usepackage{cleveref}   %% For smart ref
\usepackage{graphicx}   %% For inserting images
\graphicspath{ {./images/} }  %% configure the graphicx package
\usepackage{multirow}
\usepackage{multicol}
\usepackage{ragged2e}  %% For \raggedright
\usepackage{listings}  %% For typesetting code
\usepackage{enumitem}  %% For noitemsep and topsep
\lstdefinelanguage{smol} {
  alsoletter={ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz-!?*},
  morekeywords={defvar,deffun,lambda,set!,let,let*,letrec,begin,if,cond,else,and,or},
  morecomment=[l]{;},
  morestring=[b]",
}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines=true,
  language=smol
}

@@

%% abbrev. for RQs
\newcommand{\rqI}{What are the program-behavior misconceptions that apply to SMoL and present (even) in students with prior programming backgrounds?}
\newcommand{\rqII}{What would be a tutoring system that aims at correcting the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}

@@

\begin{document}
\maketitle
\linenumbers

@@

In the rest of this document, I first define SMoL (\Cref{s:smol}),
then presents literature review of misconceptions
(\Cref{s:background-misconceptions}) and tutoring systems
(\Cref{s:background-tutoring-systems}), then outline my research plans
for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining sections
detail my current progress.

@@

I am contacting professors from various other universities. Several of
them have expressed interests in using my instruments. So the final
results will likely include more populations.

@@

I have another source of misconceptions. I have been creating a tutor
to answer my RQ2, which I explain in details in \Cref{s:plan-rq2}. The
tutor includes many tasks that asks students to predict the result(s)
of running a program. Although these programs originate from
student-written programs (\Cref{s:find-misconceptions-with-quizius}), they
have been modified to fit tutoring purpose. Perhaps because of the
modification or because of the population variance, I have found
several unanticipated patterns of errors
(\Cref{s:misconceptions-tutor}), which suggest unknown misconceptions.

I plan to change the tutor such that it asks students' to explain
their prediction. I plan to eyeball the explanations to find new
misconceptions. This piece of data also helps me to confirm that
anticipated wrong answers do represent the expected misconceptions.

@@

The processes described in
\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
potentially introduce biases to the found misconceptions. For example,
current results find that some students think the following program
produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
\begin{lstlisting}
  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}
This checks whether $\verb|3| > \verb|n|$, but those students
presumably vocalized it as ``greater than \lstinline|3|,
\lstinline|n|?''. Although I have tried to remove programs like this
from my instruments, there is still a risk of having biase in student
errors. I plan to adjust the process descrbed in
\Cref{s:find-misconceptions-with-tutor} such that it presents programs
in multiple languages: the SMoL version, a JavaScript translation, and
a Python version.

@@

Eventually, I expect to build a table of misconceptions. Each
misconception comes with multiple programs. For each of these
programs, the wrong result that correspond to the misconception should
be commonly predicted by students from various population.
Furthermore, these students should give similar explanations for all
the wrong answers. For example, I might find a misconception called
\miscon{DefByRef} (See
\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
in \Cref{s:misconceptions-tutor} for a list of misconceptions that I
have found): when students are asked to predict the result of
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! y 0)
  x
  y
\end{lstlisting}
X\% of them responded \lstinline|12 0|; and when asked to predict
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! x 0)
  x
  y
\end{lstlisting}
Y\% of them responded \lstinline|0 12|; and these students all
explained something like ``\lstinline|y| is \lstinline|x|''.

@@

To answer this RQ, I have done some literature review on what
pedagogic techniques are helpful (\Cref{s:pedagogic-techniques}). My
current takeaway is that refutation text, case comparison, notional
machines, and language levels are helpful. I have created a tutor that
incorporates all these ideas (\Cref{s:tutor}) and have collected
data from all populations. I plan to explore more literature and to
improve the tutor if I find new helpful techniques.

I would like to first explain key aspects of the tutor
(\Cref{s:tutorials,s:interpreting-tasks}) and then my plan to
incorporate ideas of refutation text (\Cref{s:refutationt-text}), case
comparsion (\Cref{s:case-comparison}), and notional machines
(\Cref{s:notional-machine}). Finally, I present my plan to evaluate
the effectiveness of the tutor and how evaluation impacts the tutor
design (\Cref{s:evaluation}).

@@

There are many misconceptions to correct. To make the tutoring process
managable, I have design the tutor as a sequence of tutorials.
Inspired by prior research on language levels
(\Cref{s:pedagogic-techniques}), my sequence of tutorials is ordered,
and each later tutorials cover one or more language constructs
(\Cref{t:tutorials}).

Each tutorial might be delivered as multiple modules to avoid forcing
students to continuesly work for more than 30 minutes.

@@

The tutor is mostly consist of \emph{interpreting tasks}:
multiple-choice questions (MCQs) that asks students to interprete
program.

\paragraph{The MCQ Options} The options include (among other options
that I will explain shortly) the correct answer and wrong answers that
correspond to known misconceptions. Because students might conceive
wrong answers that I do not anticipate, the tutor has been providing
an ``Other'' option that, once choosen, allows students to enter
arbitrary answer.

Let's reconsider this program from \Cref{s:rq1-expected-results}
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! y 0)
  x
  y
\end{lstlisting}
In this case, the options should include the correct answer
\lstinline|12 0| and a wrong answer \lstinline|0 0|, which correspond
to the \miscon{DefByRef} misconception. If an MCQ only has a small
number of options like this example, students have a great chance (in
this case, 50\%) of just guessing the correct answer or successfully
using a process of elimination. So I have been adding additional
options that are not obviously wrong. The tutor is adding options with
an ad hoc rules. Eventually, I plan to generate options by moving
and/or copying numbers in the current options. In this case, my plan
would add \lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ
has the following options:
\begin{itemize}
  \item the correct answer,
  \item wrong answers that correspond to known misconceptions,
  \item options generated by moving and/or copying numbers in the
        current options
  \item the ``Other'' option (which, once choosen, allows students to
        enter arbitrary answer)
\end{itemize}

\paragraph{Panelty to Wrong Choices} To discourage students from
making a random choice, the tutor has been giving panelty. Designing
the strategy of assigning panelty is non-trivial:
\begin{itemize}
  \item A common strategy is to grade students by their correctness.
  However, grading by correctness is not suitable for a tutoring
  system because I do not mind students doing badly on earlier tasks
  as long as they eventually master the concepts. Actually, if
  students generally succeed in all tasks, the tutoring system is
  failing -- it does not teach anything to the students.
  \item A alternative common strategy is to grade by completion.
  However, grading by completion alone encourage students to rush
  through the tutorials and hence might encourage students to make
  random choices.
\end{itemize}
In short, I want to make students feel as follows:
\begin{itemize}
  \item They won't lose grade as long as they finish the tutorials.
  \item They should avoid giving wrong answers.
\end{itemize}
To encourage this feeling, I have been doing the following: when I
deployed the tutor in my institution, University 1, I graded by
completion\footnote{I encourage collaborators to grade by
completeness, but I have no real control over how they actually use
the tutor.}; when a student fails a task, the tutor gives an
\emph{extra} task:
\begin{enumerate}

@@

The tutor has been presenting a refutation text when a student chooses
a wrong answer that correspond to known misconception(s). Let's
reconsider the program in \Cref{s:interpreting-tasks}. If a student
answers \lstinline|0 0|, which suggests they hold the
\miscon{DefByRef} misconception, the tutor responds
\begin{quote}
  \lstinline|y| was bound to \lstinline|12| (i.e., the value of
  \lstinline|x|) rather than to \lstinline|x|. So changing the value
  of \lstinline|y| does not changes the value of \lstinline|x|.
\end{quote}
If a students gave a wrong answer that does not correspond to a
misconception (either by choosing one of the generated option or by
choosing the ``Other'' option), the tutor presents a generic
explanation. For example, if a student answers \lstinline|12 12|, the
tutor responds
\begin{quote}
  The first definition binds \lstinline|x| to \lstinline|12|. The
  second definition binds \lstinline|y| to the value of \lstinline|x|,
  which is \lstinline|12|. The \lstinline|set!| mutates the binding of
  \lstinline|y|, so \lstinline|y| is now bound to \lstinline|0|.
\end{quote}
If a student gave the correct answer, the tutor also presents the
generic explanation.

@@

The tutor provides opportunities for making case comparison: each
interpreting task presents a SMoL program and the program's result.
These program-result pairs are concrete cases of the semantics of
SMoL. To incorporate the idea of case comparsion, the tutor currently
prompts students to compare these cases and then to summarize the
rules of program behavior at the end of \emph{some} tutorials. I plan
to expand this to \emph{all} tutorials and use a consistent prompt.
The prompt should asks about the behavior of new language constructs
and their interaction with all previously introduced constructs.

The current scope tutorial uses the following prompt
\begin{quote}
  When I see a variable reference, how do I find the corresponding
  declaration, if any?
\end{quote}
This prompt has a few problems:
\begin{itemize}
  \item It is unclear to adapt it to other tutorials.
  \item Students tend to give brief response possibly because the
  questions is asking for a lot and because students are tired when
  they reach the end of a tutorial.
\end{itemize}.

@@

Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
how programs runs with notional machines (NMs) is benefitial for
students. I have created a web-based notional machine
\begin{center}
  \href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}
\end{center}

The notional machine can be configurated with URL parameters. For
example,
\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
link} opens the NM with a program embeded.

@@

To fully answer my research questions, I should also argue that my
tutor is effective at correcting misconceptions. In this secton, I
describe how I want to refine the design of the tutor to help me make
the argument, what data I want to collect, and what (statistical)
analysis do I want to perform on the data.

I can infer whether students hold a misconception by analysing their
responses to interpreting tasks. There are three possible responses to
an interpreting task:
\begin{enumerate}
  \item They give the correct answer.
  \item They give a wrong answer that correspond to some
        misconceptions.
  \item They give a wrong answer that do not correspond to any (known)
        misconceptions.
\end{enumerate}
If a student gives a wrong answer corresponds to exactly one
misconception \emph{and} an explanation that matches the
misconception, I deduce that the student holds the misconception when
doing the task.

@@

If the \textbf{TaskSpecific} model best explains the data, I deduce
that some programs are more confusing and that I need further
investigation into the relationship between the tasks and
misconceptions. If the \textbf{PositionSpecific} model best explains
\emph{and} students perform better in later tasks, I deduce that the
tutor seeems to be correcting misconceptions. If studnets perform
worse in later taks, the tutor is potentially doing harm.

To know students well, I would like to have as many tasks per
misconception as possible. But in practice, the number of tasks is
subject to the following limitations:
\begin{enumerate}
  \item I don't want the length of any tutorial section to exceed 30
  minutes.
  \item Programs shouldn't be too long. Some misconceptions are really
  similar and as a consequence have only a small number of short
  programs to tell them apart.
\end{enumerate}

The current tutor is not in the final shape yet. In particular, many
of the wrong answers correspond to more than one misconceptions; some
misconceptions are represented by only one question; the order of
interpreting tasks are not randomized. I plan to tackles the first two
differences by adding and removing programs. Adding randomization
requires me to carefully check the dependencies between current
explanations and other texts in the tutor and resolve those
dependencies if any.

@@

\Cref{f:smol-tutor-example} illustrates an interpreting task as
descrbed in \Cref{s:interpreting-tasks}. The Tutor also asked students
to perform some other activities, like showing programs and asking for
their heap content, which I do not cover in this paper.

@@

By the end of Fall 2023, I would like to improve the SMoL Tutor:
\begin{enumerate}
  \item Ask students to explained when they gave a wrong answer in a
        interepreting question. With student explanations, I check whether
        choosing the wrong answer indicates holding the corresponding misconception(s).
  \item Adjust the collection of interpreting questions such that
        every misconception has at least two questions that aims at
        correcting the misconception.
  \item Partially randomize the order of questions.
\end{enumerate}

@@

During Fall 2024, I plan to collect a new round of data from Brown. I
expect this datasets and the datasets from Spring 2024 would give a
more rigorously evaluation of the effectiveness of SMoL Tutor. The
improvements to be made in Fall 2023 should resolve many of the
threats to validity. In this semester I also want to write a new paper
about the SMoL Tutor.

>>>>

% \usepackage[switch]{lineno}
% \renewcommand\thelinenumber{\textcolor{red}{\arabic{linenumber}}}
\usepackage{xcolor}
\usepackage[utf8]{inputenc} % Use UTF-8 encoding
\usepackage[T1]{fontenc} % Use the T1 font encoding
\usepackage{geometry} % Adjust page margins
\usepackage{amsmath, amssymb} % For math symbols and equations
\usepackage{graphicx} % For including images
\usepackage{titlesec} % For customizing section titles
% \usepackage[style=authoryear]{biblatex}  % For managing references
\usepackage[round, square]{natbib}
\bibliographystyle{plainnat}
\usepackage{hyperref}
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{tabularx}   %% For tables with growable column
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\usepackage{cleveref}   %% For smart ref
\usepackage{graphicx}   %% For inserting images
\graphicspath{ {./images/} }  %% configure the graphicx package
\usepackage{multirow}
\usepackage{multicol}
\usepackage{ragged2e}  %% For \raggedright
\usepackage{listings}  %% For typesetting code
\usepackage{enumitem}  %% For noitemsep and topsep
\lstdefinelanguage{smol} {
  alsoletter={ABCDEFGHIJKLMNOPQRSTUVWXYZabcdefghijklmnopqrstuvwxyz-!?*},
  morekeywords={defvar,deffun,lambda,set!,let,let*,letrec,begin,if,cond,else,and,or},
  morecomment=[l]{;},
  morestring=[b]",
}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines=true,
  language=smol
}

@@

%% abbrev. for RQs
\newcommand{\rqI}{What program-behavior misconceptions apply to SMoL and present (even) in students with prior programming backgrounds?}
\newcommand{\rqII}{What would be a tutoring system that aims to correct the misconceptions, draws on cognitive, educational, or psychological concepts, and seems to be effective?}

@@

\begin{document}
\maketitle
% \linenumbers

@@

In the rest of this document, I first define SMoL (\Cref{s:smol}),
then present a literature review of misconceptions
(\Cref{s:background-misconceptions}) and tutoring systems
(\Cref{s:background-tutoring-systems}), and then outline my research
plans for the RQs (\Cref{s:plan-rq1,s:plan-rq2}). The remaining
sections detail my current progress.

@@

I am contacting professors from various other universities. Several of
them have expressed interest in using my instruments. So, the final
results will likely include more populations.

@@

I have another source of misconceptions. I have been creating a tutor
to answer my RQ2, which I explain in detail in \Cref{s:plan-rq2}. The
tutor includes many tasks that ask students to predict the result(s)
of running a program. Although these programs originate from
student-written programs (\Cref{s:find-misconceptions-with-quizius}), they
have been modified to fit tutoring purposes. Perhaps because of the
modification or because of the population variance, I have found
several unanticipated patterns of errors
(\Cref{s:misconceptions-tutor}), which suggest unknown misconceptions.

I plan to change the tutor such that it asks students to explain
their prediction. I plan to eyeball the explanations to find new
misconceptions. This piece of data also helps me confirm that
anticipated wrong answers represent the expected misconceptions.

@@

The processes described in
\Cref{s:find-misconceptions-with-quizius,s:find-misconceptions-with-tutor}
have been conducted with SMoL's Lispy syntax (\Cref{s:smol}). This can
potentially introduce biases to the found misconceptions. For example,
current results find that some students think the following program
produces \lstinline|'(4 5)| rather than \lstinline|'(1 2)|.
\begin{lstlisting}
  (filter (lambda (n) (> 3 n)) '(1 2 3 4 5))
\end{lstlisting}
This checks whether $\verb|3| > \verb|n|$, but those students
presumably vocalized it as ``greater than \lstinline|3|,
\lstinline|n|?''. Although I have tried to remove programs like this
from my instruments, there is still a risk of having biase in student
errors. I plan to adjust the process described in
\Cref{s:find-misconceptions-with-tutor} so that it presents programs
in multiple languages: the SMoL version, a JavaScript translation, and
a Python version.

@@

Eventually, I expect to build a table of misconceptions. Each
misconception comes with multiple programs. For each of these
programs, the wrong result that corresponds to the misconception should
be commonly predicted by students from various population.
Furthermore, these students should give similar explanations for all
the wrong answers. For example, I might find a misconception called
\miscon{DefByRef} (See
\Cref{t:tutor-grounded-misconceptions-1,t:tutor-grounded-misconceptions-2,t:tutor-surmised-misconceptions}
in \Cref{s:misconceptions-tutor} for a list of misconceptions that I
have found): when students are asked to predict the result of
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! y 0)
  x
  y
\end{lstlisting}
X\% of them responded \lstinline|12 0|; when asked to predict
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! x 0)
  x
  y
\end{lstlisting}
Y\% of them responded \lstinline|0 12|; and these students all
explained something like ``\lstinline|y| is \lstinline|x|''.

@@

To answer this RQ, I have done some literature review on what
pedagogic techniques are helpful (\Cref{s:pedagogic-techniques}). My
current takeaway is that refutation text, case comparison, notional
machines, and language levels are helpful. I have created a tutor that
incorporates all these ideas (\Cref{s:tutor}) and have collected
data from all populations. I plan to explore more literature and
improve the tutor if I find new helpful techniques.

I would like to first explain key aspects of the tutor
(\Cref{s:tutorials,s:interpreting-tasks}) and then my plan to
incorporate ideas of refutation text (\Cref{s:refutationt-text}), case
comparison (\Cref{s:case-comparison}), and notional machines
(\Cref{s:notional-machine}). Finally, I present my plan to evaluate
the effectiveness of the tutor and how evaluation impacts the tutor
design (\Cref{s:evaluation}).

@@

There are many misconceptions to correct. To make the tutoring process
manageable, I have designed the tutor as a sequence of tutorials.
Inspired by prior research on language levels
(\Cref{s:pedagogic-techniques}), my sequence of tutorials is ordered,
and each tutorial covers one or more new language constructs
(\Cref{t:tutorials}).

Each tutorial might be delivered as multiple modules to avoid forcing
students to continuously work for more than 30 minutes.

@@

The tutor is mostly consisted of \emph{interpreting tasks}:
multiple-choice questions (MCQs) that asks students to interpret
program.

\paragraph{The MCQ Options} The options include (among other options
that I will explain shortly) the correct answer and wrong answers that
correspond to known misconceptions. Because students might conceive
wrong answers that I do not anticipate, the tutor has been providing
an ``Other'' option that, once chosen, allows students to enter
arbitrary answers.

Let's reconsider this program from \Cref{s:rq1-expected-results}
\begin{lstlisting}
  (defvar x 12)
  (defvar y x)
  (set! y 0)
  x
  y
\end{lstlisting}
In this case, the options should include the correct answer
\lstinline|12 0| and a wrong answer \lstinline|0 0|, which corresponds
to the \miscon{DefByRef} misconception. If an MCQ only has few options
like this example, students have a great chance (in this case, 50\%)
of just guessing the correct answer or successfully using a process of
elimination. So I have been adding additional options that are not
obviously wrong. The tutor is adding options with an ad hoc rule.
Eventually, I plan to generate options by moving and/or copying
numbers in the current options. In this case, my plan would add
\lstinline|12 12| and \lstinline|0 12|. In summary, each MCQ has the
following options:
\begin{itemize}
  \item the correct answer,
  \item wrong answers that correspond to known misconceptions,
  \item options generated by moving and/or copying numbers in the
        current options
  \item the ``Other'' option (which, once chosen, allows students to
        enter arbitrary answers)
\end{itemize}

\paragraph{Penalty to Wrong Choices} To discourage students from
making a random choice, the tutor has been giving penalty. Designing
the strategy of assigning penalty is non-trivial:
\begin{itemize}
  \item A common strategy is to grade students by their correctness.
  However, grading by correctness is not suitable for a tutoring
  system because I do not mind students doing badly on earlier tasks
  as long as they eventually master the concepts. Actually, if
  students generally succeed in all tasks, the tutoring system is
  failing -- it does not teach anything to the students.
  \item An alternative common strategy is to grade by completion.
  However, grading by completion alone encourages students to rush
  through the tutorials and hence might encourage students to make
  random choices.
\end{itemize}
In short, I want to make students feel as follows:
\begin{itemize}
  \item They won't lose grades as long as they finish the tutorials.
  \item They should avoid giving wrong answers.
\end{itemize}
To encourage this feeling, I have been doing the following: when I
deployed the tutor in my institution, University 1, I graded by
completion\footnote{I encourage collaborators to grade by
completeness, but I have no real control over how they actually use
the tutor.}; when a student fails a task, the tutor gives an
\emph{extra} task:
\begin{enumerate}

@@

The tutor has been presenting a refutation text when a student chooses
a wrong answer that corresponds to known misconception(s). Let's
reconsider the program in \Cref{s:interpreting-tasks}. If a student
answers \lstinline|0 0|, which suggests they hold the
\miscon{DefByRef} misconception, the tutor responds
\begin{quote}
  \lstinline|y| was bound to \lstinline|12| (i.e., the value of
  \lstinline|x|) rather than to \lstinline|x|. So changing the value
  of \lstinline|y| does not change the value of \lstinline|x|.
\end{quote}
If a student gave a wrong answer that does not correspond to a
misconception (either by choosing one of the generated option or by
choosing the ``Other'' option), the tutor presents a generic
explanation. For example, if a student answers \lstinline|12 12|, the
tutor responds
\begin{quote}
  The first definition binds \lstinline|x| to \lstinline|12|. The
  second definition binds \lstinline|y| to the value of \lstinline|x|,
  which is \lstinline|12|. The \lstinline|set!| mutates the binding of
  \lstinline|y|, so \lstinline|y| is now bound to \lstinline|0|.
\end{quote}
If a student gave the correct answer, the tutor also presents the
generic explanation.

@@

The tutor provides opportunities for making case comparisons: each
interpreting task presents a SMoL program and the program's result.
These program-result pairs are concrete cases of the semantics of
SMoL. To incorporate the idea of case comparison, the tutor currently
prompts students to compare these cases and then to summarize the
rules of program behavior at the end of \emph{some} tutorials. I plan
to expand this to \emph{all} tutorials and use a consistent prompt.
The prompt should ask about the behavior of new language constructs
and their interaction with all previously introduced constructs.

The current scope tutorial uses the following prompt:
\begin{quote}
  When I see a variable reference, how do I find the corresponding
  declaration, if any?
\end{quote}
This prompt has a few problems:
\begin{itemize}
  \item It is unclear to adapt it to other tutorials.
  \item Students tend to give brief responses possibly because the
  questions is asking for a lot and because students are tired when
  they reach the end of a tutorial.
\end{itemize}

@@

Prior research argues (\Cref{s:pedagogic-techniques})that illustrating
how programs runs with notional machines (NMs) is beneficial for
students. I have created a web-based notional machine
\begin{center}
  \href{https://lukuangchen.github.io/stacker-2023}{https://lukuangchen.github.io/stacker-2023}
\end{center}

The notional machine can be configured with URL parameters. For
example,
\href{https://lukuangchen.github.io/stacker-2023/?program=\%28defvar+x+12\%29\%0A\%28defvar+y+x\%29\%0A\%28set\%21+y+0\%29\%0Ax\%0Ay}{this
link} opens the NM with a program embedded.

@@

To fully answer my research questions, I should also argue that my
tutor is effective at correcting misconceptions. In this section, I
describe how I want to refine the design of the tutor to help me make
the argument, what data I want to collect, and what (statistical)
analysis do I want to perform on the data.

I can infer whether students hold a misconception by analyzing their
responses to interpreting tasks. There are three possible responses to
an interpreting task:
\begin{enumerate}
  \item They give the correct answer.
  \item They give a wrong answer that correspond to some
        misconceptions.
  \item They give a wrong answer that do not correspond to any (known)
        misconceptions.
\end{enumerate}
If a student gives a wrong answer corresponds to exactly one
misconception \emph{and} an explanation that matches the
misconception, I deduce that the student holds the misconception when
doing the task.

@@

If the \textbf{TaskSpecific} model best explains the data, I deduce
that some programs are more confusing and that I need further
investigation into the relationship between the tasks and
misconceptions. If the \textbf{PositionSpecific} model best explains
\emph{and} students perform better in later tasks, I deduce that the
tutor seems to be correcting misconceptions. If students perform
worse in later tasks, the tutor is potentially doing harm.

To know students well, I would like to have as many tasks per
misconception as possible. But in practice, the number of tasks is
subject to the following limitations:
\begin{enumerate}
  \item I don't want the length of any tutorial section to exceed 30
  minutes.
  \item Programs shouldn't be too long. Some misconceptions are really
  similar and as a consequence have few short programs to tell them
  apart.
\end{enumerate}

The current tutor is not in the final shape yet. In particular, many
of the wrong answers correspond to more than one misconception; some
misconceptions are represented by only one question; the order of
interpreting tasks are not randomized. I plan to tackle the first two
differences by adding and removing programs. Adding randomization
requires me to carefully check the dependencies between current
explanations and other texts in the tutor and resolve those
dependencies if any.

@@

\Cref{f:smol-tutor-example} illustrates an interpreting task as
described in \Cref{s:interpreting-tasks}. The Tutor also asked students
to perform some other activities, like showing programs and asking for
their heap content, which I do not cover in this paper.

@@

By the end of Fall 2023, I would like to improve the SMoL Tutor:
\begin{enumerate}
  \item Ask students to explain when they gave a wrong answer in an
        interpreting question. With student explanations, I check whether
        choosing the wrong answer indicates holding the corresponding misconception(s).
  \item Adjust the collection of interpreting questions such that
        every misconception has at least two questions that aims at
        correcting the misconception.
  \item Partially randomize the order of questions.
\end{enumerate}

@@

During Fall 2024, I plan to collect a new round of data from Brown. I
expect these datasets and the datasets from Spring 2024 would give a
more rigorously evaluation of the effectiveness of SMoL Tutor. The
improvements to be made in Fall 2023 should resolve many of the
threats to validity. In this semester I also want to write a new paper
about the SMoL Tutor.
